---
title: "Assignment – Tidying and Transforming Data"
author: "Mathew Katz"
date: "2022-10-01"
output:
  pdf_document: default
  html_document: default
---
Read in necessary libraries: 
```{r} 
library(tidyr)  
library(dplyr)
library(stringr)
library(ggplot2)
```
Read in csv and look at it:
```{r}
flights <- read.csv(file = 'flight_tidying.csv') 
flights 
```
Look at structure of csv:
```{r}
str(flights) 
```
Change columnn names to 'better' names
```{r}
flights <- flights %>% 
    rename('Airline' = 1, 'Flight.Status' = 2)
flights
```
Remove NAs and empty rows in csv:
```{r}
flights <- flights %>% #
            filter(! is.na(Flight.Status) &  
                     str_length(Flight.Status) > 0) 
flights
```
Turn csv into dataframe:
```{r}
df <- data.frame(flights)
head(df)
```
Add missing airline information:
```{r}
df[2, "Airline"] <- "Alaska"
df[4, "Airline"] <- "AM WEST"
df
```
Lengthen data by increasing the number of rows and decreasing the number of columns:
```{r}
df <- df %>%
        pivot_longer(!c("Airline", "Flight.Status"),
                     names_to = "Destination",               
                     values_to = "Count")      
df
```
Write to CSV:
```{r}
write.csv(df, "./clean_flight_tidying.csv", row.names=FALSE)
```
Create two new dataframes of delayed and non time flights to graph them:
```{r}
delayed_flights <- df %>%
                    filter(df$Flight.Status == "delayed")
delayed_flights
```
```{r}
on_time_flights <- df %>%
                    filter(df$Flight.Status == "on time")
on_time_flights
```
Graph Delayed Flights:
```{r}
 delayed_bar_graph <- ggplot(data=delayed_flights, aes(x=Destination, y=Count, fill=Airline))
    delayed_bar_graph <- delayed_bar_graph +  ggtitle('Delayed Flights') +
            geom_bar(stat="identity", position=position_dodge()) 
delayed_bar_graph
```
AM WEST has more delayed flights than Alaska in ever city but Seattle. 

Graph On-Time Flights:
```{r}
 ontime_bar_graph <- ggplot(data=on_time_flights, aes(x=Destination, y=Count, fill=Airline))
    ontime_bar_graph <- ontime_bar_graph +  ggtitle('On-Time Flights') +
            geom_bar(stat="identity", position=position_dodge()) 
ontime_bar_graph
```
AM WEST has more on-time flights than Alaska in Los Angeles, Phoenix, and San Diego. 
Alaska has more on-time flights than AM WEST in San Francisco and Seattle. 



















