---
title: "SentAnalysisData607"
author: "Mathew Katz"
date: "2022-11-06"
output:
  pdf_document: default
  html_document: default
---

Firstly I am going to post a bunch of code from Text Mining with R, Chapter 2 (Sentiment Analysis.) 
{https://www.tidytextmining.com/sentiment.html}

```{r}
library(tidytext)
get_sentiments("afinn")
```

```{r}
get_sentiments("bing")
```

```{r}
get_sentiments("nrc")
```

```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r}
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)
```

```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```

```{r}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

```{r}
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
```

```{r}
p_and_p_sentences$sentence[2]
```


```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```


```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

Let's extend the code in two ways:
Work with a different corpus of our choosing, and
Incorporate at least one additional sentiment lexicon (possibly from another R package that I’ve found through research).

```{r}
library('readr')
scripts <- read_csv("RickAndMortyScripts.csv")
```


```{r}
nrc <- get_sentiments("nrc")
bing <- get_sentiments("bing")
afinn <- get_sentiments("afinn")
```

```{r}
scripts = scripts %>% rename(Index = "index",
                   Season.No = "season no.",
                   Episode.No = "episode no.",
                   Episode.Name = "episode name",
                   Character.Name = "name",
                   Dialog = "line")
```

```{r}
# Head of the table
head(scripts, 4)

# Tail of the table
tail(scripts, 4)

# Summary
summary(scripts)
```

Clean Corpus Function:
This predefined function is going to clean the text from:

the punctuation - removePunctuation
extra white space - stripWhitespace
transforms to lower case - tolower
stopwords (common words that should be ignored) - stopwords
numbers - removeNumbers
```{r}
cleanCorpus <- function(text){
  # punctuation, whitespace, lowercase, numbers
  text.tmp <- tm_map(text, removePunctuation)
  text.tmp <- tm_map(text.tmp, stripWhitespace)
  text.tmp <- tm_map(text.tmp, content_transformer(tolower))
  text.tmp <- tm_map(text.tmp, removeNumbers)
  
  # removes stopwords
  stopwords_remove <- c(stopwords("en"), c("thats","weve","hes","theres","ive","im",
                                                "will","can","cant","dont","youve","us",
                                                "youre","youll","theyre","whats","didnt"))
  text.tmp <- tm_map(text.tmp, removeWords, stopwords_remove)

  return(text.tmp)
}
```
These predefined functions will process the text depending on the case:

Unigrams take only 1 word at a time
Bigrams take 2 sequential words at a time
Trigrams (you guessed) take 3 sequential words at a time
Eg. text: "come on morty"

Unigram: "come", "on", "morty"
Bigram: "come on", "on morty"
Trigram: "come on morty"
Term Document Matrix: it's a mathematical matrix that describes the frequency of terms that occur in a collection of documents. More simply put, is a matrix that has on:

rows - words that can be found in the analysed documents
columns - the documents in order
values - the frequency of each word in each document

Unigram:
```{r}
frequentTerms <- function(text){
  
  # create the matrix
  s.cor <- VCorpus(VectorSource(text))
  s.cor.cl <- cleanCorpus(s.cor)
  s.tdm <- TermDocumentMatrix(s.cor.cl)
  s.tdm <- removeSparseTerms(s.tdm, 0.999)
  m <- as.matrix(s.tdm)
  word_freqs <- sort(rowSums(m), decreasing = T)
  
  # change to dataframe
  dm <- data.frame(word=names(word_freqs), freq=word_freqs)
  
  return(dm)
}
```
Bigram:
```{r}
# Bigram tokenizer
tokenizer_2 <- function(x){
  NGramTokenizer(x, Weka_control(min=2, max=2))
}

# Bigram function 
frequentBigrams <- function(text){

  s.cor <- VCorpus(VectorSource(text))
  s.cor.cl <- cleanCorpus(s.cor)
  s.tdm <- TermDocumentMatrix(s.cor.cl, control=list(tokenize=tokenizer_2))
  s.tdm <- removeSparseTerms(s.tdm, 0.999)
  m <- as.matrix(s.tdm)
  word_freqs <- sort(rowSums(m), decreasing=T)
  dm <- data.frame(word=names(word_freqs), freq=word_freqs)
  
  return(dm)
}
```
Trigram:
```{r}
# Trigram tokenizer
tokenizer_3 <- function(x){
  NGramTokenizer(x, Weka_control(min=3, max=3))
}

# Trigram function 
frequentTrigrams <- function(text){

  s.cor <- VCorpus(VectorSource(text))
  s.cor.cl <- cleanCorpus(s.cor)
  s.tdm <- TermDocumentMatrix(s.cor.cl, control=list(tokenize=tokenizer_3))
  s.tdm <- removeSparseTerms(s.tdm, 0.999)
  m <- as.matrix(s.tdm)
  word_freqs <- sort(rowSums(m), decreasing=T)
  dm <- data.frame(word=names(word_freqs), freq=word_freqs)
  
  return(dm)
}
```
Bing Lexicon cathegorizes the words into positives and negatives.

To be able to do so in our data, first we make a dataframe that splits all the words in 1 dialogue onto rows. Afterwards, we can join our data with the lexicon, leaving us with a beautiful classification of our words.


```{r}
# Creating our tokens
tokens <- scripts %>% 
  mutate(dialogue = as.character(scripts$Dialog)) %>% 
  unnest_tokens(word, dialogue)

tokens %>% head(5) %>% select(Character.Name, word)
```

```{r}
tokens %>% 
  # append the bing sentiment and prepare the data
  inner_join(bing, "word") %>%
  count(word, sentiment, sort=T) %>% 
  acast(word ~ sentiment, value.var = "n", fill=0) %>% 
  
  # wordcloud
  comparison.cloud(colors=c("#991D1D", "#327CDE"), max.words = 100)
```
How is the overall mood in Rick & Morty?

The nrc lexicon cathegorizes the words in 10 moods:

positive
negative
anger
anticipation
disgust
fear
joy
sadness
surprise
trust
Let's look at how these sentiments rank in out data:
```{r}
sentiments <- tokens %>% 
  inner_join(nrc, "word") %>%
  count(sentiment, sort=T)

sentiments
```
Afinn Lexicon ranks every word from -5 to 5, where:

-5 being the most negative
+5 being the most positive
```{r}
tokens %>% 
  # Count how many word per value
  inner_join(afinn, "word") %>% 
  count(value, sort=T) %>%
  
  # Plot
  ggplot(aes(x=value, y=n)) +
  geom_bar(stat="identity", aes(fill=n), show.legend = F, width = 0.5) +
  geom_label(aes(label=n)) +
  scale_fill_gradient(low="#85C1E9", high="#3498DB") +
  scale_x_continuous(breaks=seq(-5, 5, 1)) +
  labs(x="Score", y="Frequency", title="Word count distribution over intensity of sentiment: Neg -> Pos") +
  theme_bw()
```

```{r}
tokens %>% 
  # by word and value count number of occurences
  inner_join(afinn, "word") %>% 
  count(word, value, sort=T) %>% 
  mutate(contribution = n * value,
         sentiment = ifelse(contribution<=0, "Negative", "Positive")) %>% #another variable
  arrange(desc(abs(contribution))) %>% 
  head(20) %>% 
  
  # plot
  ggplot(aes(x=reorder(word, contribution), y=contribution, fill=sentiment)) +
  geom_col(aes(fill=sentiment), show.legend = F) +
  labs(x="Word", y="Contribution", title="Words with biggest contributions in positive/negative moods") +
  coord_flip() +
  scale_fill_manual(values=c("#FA8072", "#08439A")) + 
  theme_bw()

```


```{r}
library("stopwords")

# Create a dataframe with stopwords
stopwords_script <- tibble(word = c(stopwords("en"), c("thats","weve","hes","theres","ive","im",
                                                           "will","can","cant","dont","youve","us",
                                                           "youre","youll","theyre","whats","didnt")))
print(stopwords_script)


# Create the dataframe of tokens
scripts %>% 
  mutate(dialogue = as.character(scripts$Dialog)) %>% 
  filter(Character.Name %in% c("Rick","Morty","Beth","Jerry","Summer")) %>% 
  
  # removes stopwords
  unnest_tokens(word, dialogue) %>%
  anti_join(stopwords_script, by="word") %>%
  
  # top N frequent words per character
  count(Character.Name, word) %>% 
  group_by(Character.Name) %>% 
  arrange(desc(n)) %>% 
  slice(1:10) %>% 
  
  mutate(word2 = factor(paste(word, Character.Name, sep="__"),
                        levels = rev(paste(word, Character.Name, sep="__"))))
```

Rick and Morty's script made this project very enjoyable.  